---
title: "Results"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Community elevation analysis
Dissimilarity in dipterocarp community compositions reflected by colour variations across the 160 ha plot, extracted from axis 1 of a non-metric multi-dimensional scaling analysis of 12 elevational bands. 

```{r warning=FALSE, echo=FALSE, results='hide', message=FALSE, fig.align='center', fig.dim=c(8, 4)}
fig1c <- source('./src/analysis/community-elevation-analysis/community-elevation.graph.R')$value
fig1c$plot
```


## Species distibution modelling

The elevation distribution of adult dipterocarps within a 160 ha forest plot using the Integrated Nested Laplace Approximation. The 95% credible intervals (n=5000) were extracted from the posterior distribution. 

```{r warning=FALSE,echo=FALSE, fig.align='center', fig.dim=c(7, 4)}
fig2a <- source('./src/analysis/inla-spde-species-distributions/species-distribution.graph.R')$value
fig2a$plot
```



## Species water innundation sensitivity 

Species differ in seedling sensitivity to water inundation along micro-topographical gradients within the alluvial zone. Points are ordered by the most likely elevation the adults are found, and grouped by below and above average wood densities for these species. 95% confidence intervals are bootstrapped (n = 5000).

```{r warning=FALSE,echo=FALSE, fig.align='center', fig.dim=c(8, 3)}
fig2b <- source('./src/analysis/sepilok-field-experiment/mortality-exp.graph.R')$value
fig2b$plot
```

Species varied in their tolerances to water inundation (ΔAICsp:inundation = 4.96, Fig. 2b & S1, Table S1). At one extreme, seedlings of Shorea seminis showed no difference in the probability of mortality between inundated and non-inundated areas. Shorea beccariana, representing the other end of the response gradient, had an absolute risk increase in mortality of 40.9% when inundated (Fig. 2b & S1). Seedling mortality was not affected by soil pH, soil texture, light, or flash floods (178 mm of rainfall in 6 hours; see Methods), nor did inclusion of any of these predictors remove the relationship between seedling mortality and inundation. We refer to the difference in seedling mortality (Δp(mortality)) with respect to water inundation as “inundation sensitivity”. We observed that the species with the highest adult wood densities were more resistant to inundation (Fig. 2b & S1).


## Species distribution explained by the additive effect of innundation sensitvity and wood density
Multivariate regression of inundation sensitivity (a) and wood density (b) predicting the estimated adult elevation distribution (m). Lines and 95% Confidence intervals were predicted from the model. Red points indicate the raw data, grey points are the partial residuals, and red dashed lines indicate the univariate analysis. Sixteen species were included in the analysis. Percentage variance for each variable from an ANOVA test.

```{r warning=FALSE, echo=FALSE, results='hide', message=FALSE, fig.align='center', fig.dim=c(8, 4)}
gg_theme <- source('./src/utils/gg-theme.graph.R')$value
fig3a <- source('./src/analysis/elevation-predicted-by-density+inundation/riskratio.graph.R')$value
fig3b <- source('./src/analysis/elevation-predicted-by-density+inundation/density.graph.R')$value
ggpubr::ggarrange(
    fig3a$plot,
    fig3b$plot,
    ncol = 2, 
    nrow = 1, 
    labels = c('a', 'b'),
    font.label = list(size = 14, face = "bold", family = gg_theme$font_family) 
  )
```

Inundation sensitivity explained 37.3% of interspecific variation in species elevation distribution (βinundation sensitivity = 148.39, 95% CI = 51.53 – 245.24, Fig. 2a), indicating that tolerant individuals occur where areas with inundation episodes. When controlling for inundation, our model predicts that a positive relationship between wood density and elevation (βwood density =106.63, 95% CI = 7.71 – 205.55, anova = 18.5%, Fig. 2b). This suggests that species with high wood density as adults coexist on the alluvial plain and ridge tops partly because of water inundation episodes and droughts. 


## Ontogenic stages wood density 
```{r warning=FALSE, echo=FALSE, results='hide', message=FALSE, fig.align='center', fig.dim=c(4, 4)}
fig3c <- source('./src/analysis/ontogenic-wood-density-change/ontogenic-density.graph.R')$value # the triangle

fig3c$plot
```
We find that wood density is not conserved through ontogenic stages (Fig 3c), and therefore it would be premature to conclude wood density provides tolerance to seedling. 


## Wood density changes due to water inundation 
```{r warning=FALSE, echo=FALSE, results='hide', message=FALSE, fig.align='center', fig.dim=c(4, 4)}
fig3d <- source('./src/analysis/nursery-experiment/final-wood-density.graph.R')$value
fig3d$plot # plot 

# calc typical percent increase
min_density <- min(fig3d$analysis$preds_treat$final_wood_density)
max_density <- max(fig3d$analysis$preds_treat$final_wood_density)
percent_increase <- round((max_density - min_density) / min_density * 100, 1)

```
... In a nursery experiment we found water inundation to increase seedling wood densities (βlog(inundation frequency + 1) = `r round(fig3d$analysis$coef['log(treatment + 1)', 'Estimate'], 4)`, 95% CI = `r round(fig3d$analysis$coef['log(treatment + 1)', "CI025"], 4)` – `r round(fig3d$analysis$coef['log(treatment + 1)', "CI975"], 4)`, Fig. 3d). In the most severe inundation regime, the seedling wood density increased by `r percent_increase`% (from `r round(min_density, 3)` to `r round(max_density, 3)` g cm-3). This result indicates that water inundation has the potential to induce high wood density among dipterocarps. Therefore adult wood densities predicting species distributions may reflect life history and a capacity to grow in a particular set of environment conditions...